---
title: "Exam 1"
author: "STAT 211 - 509"
date: "10-4-2018"
output: pdf_document
params:
  seed: 1
  key: TRUE
subtitle: "Version `r params$seed`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 4)

# VERSION A
set.seed(params$seed)

# VERSION B
# set.seed(2)

answer <- function(ans, choices) {
  paste0("SOLUTION: ", letters[which(ans == choices)])
}

answer_list <- function(ans, choices) {
  paste0("SOLUTION: ", letters[which(sapply(choices, function(e) all(e == ans)))])
}
```


# Name: _____________________

Please sign the following pledge and read all instructions carefully before starting the exam.

Pledge: I have neither given nor received any unauthorized aid in completing this exam, and I have conducted myself within the guidelines of the University Honor Code.
Signature: ________________________

# Instructions

+ This test has 15 multiple choice problems. Record all answers on your large 
  Scantron with the appropriate pencil.
+ Total time is 75 minutes (12:45 P.M to 2:00 P.M.).
+ You are permitted one 8.5in by 11in cheatsheet and a calculator. No other 
  resources (phones, laptop, tablet, etc.) are allowed. 
+ If you are wearing a hat and/or smartwatch, please remove them while taking the
  test.
+ There is no penalty for incorrect answers; have an answer for every question.
+ When you are done, turn in your cheatsheet and exam booklet along with the Scantron.

\newpage

```{r echo=FALSE, results='hide'}
q1_soln <- "a + 4b"
q1_vec <- sample(c(
  q1_soln,
  "a + b",
  "a + 8b",
  "2a + 4b",
  "2a + b"
))

q2_soln <- "$a^2 + 2b^2$"
q2_vec <- sample(c(
  q2_soln,
  "$a^2 + b^2$",
  "$a^2 + 4b^2$",
  "$2b^2$",
  "$a^2$" 
))

q3_soln <- "$a^2 + 2b^2 + (a + 4b)^2$"
q3_vec <- sample(c(
  q3_soln,
  "$(a+4b)^2$",
  "$a^2 + 2b^2 - (a + 4b)^2$",
  "$a^2 + 2b^2 + a + 4b$",
  "$a^2 + 2b^2 - a - 4b$"
))
```

Let $X \sim Poisson(1)$ and $Y \sim Binomial(8, 1/2)$. Assume that $X$ and $Y$
are independent. Answer the following **three** questions:

## Question 1

What is $E[aX + bY]$?
 
\begin{enumerate}
  \item[a)] `r q1_vec[1]`
  \item[b)] `r q1_vec[2]`
  \item[c)] `r q1_vec[3]`
  \item[d)] `r q1_vec[4]`
  \item[e)] `r q1_vec[5]`
\end{enumerate}

`r if (params$key) answer(q1_soln, q1_vec)`

## Question 2

What is $Var(aX + bY)$?

\begin{enumerate}
  \item[a)] `r q2_vec[1]`
  \item[b)] `r q2_vec[2]`
  \item[c)] `r q2_vec[3]`
  \item[d)] `r q2_vec[4]`
  \item[e)] `r q2_vec[5]`
\end{enumerate}

`r if (params$key) answer(q2_soln, q2_vec)`

## Question 3

What is $E[(aX + bY)^2]$?

\begin{enumerate}
  \item[a)] `r q3_vec[1]`
  \item[b)] `r q3_vec[2]`
  \item[c)] `r q3_vec[3]`
  \item[d)] `r q3_vec[4]`
  \item[e)] `r q3_vec[5]`
\end{enumerate}

`r if (params$key) answer(q3_soln, q3_vec)`

*****

\newpage

## Question 4

```{r echo=FALSE, results='hide'}
p_L <- sample(seq(.1, .4), size = 1)
p_T <- sample(seq(.3, .5), size = 1)
p_LUT <- p_L + p_T

q4_vec <- sample(list(
  p_LUT,
  p_L * p_T,
  (1 - p_L) * p_T,
  "Can't answer without knowing $P(A \\cap B)$",
  p_LUT + .05
))
```

Let $L$ be the event that your bus is late and $T$ be the event that it is on time.
Suppose $P(L) = `r p_L`$ and $P(T) = `r p_T`$. What is $P(L \cup T)$?

\begin{enumerate}
  \item[a)] `r q4_vec[1]`
  \item[b)] `r q4_vec[2]`
  \item[c)] `r q4_vec[3]`
  \item[d)] `r q4_vec[4]`
  \item[e)] `r q4_vec[5]`
\end{enumerate}

`r if (params$key) answer(p_LUT, q4_vec)`

*****

## Question 5

```{r echo=FALSE, results='hide'}
q5_soln <- "P(B)"
q5_vec <- sample(list(
  q5_soln,
  "P(A)",
  1,
  0,
  "P(A) + P(B)"
))
```

Suppose for two events $A$ and $B$ we have that $B \subset A$, meaning $B$ is completely contained 
in $A$. This means every outcome in $B$ is also in $A$. Then $P(A \cap B)$ is:

*Hint: draw a Venn diagram. For an example of an event completely contained in 
another, let $P$ be the event you pass this exam and $A$ be the event you get an A.
$A \subset P$, since all the scores you get that lead to an A also lead to you passing.*

\begin{enumerate}
  \item[a)] `r q5_vec[1]`
  \item[b)] `r q5_vec[2]`
  \item[c)] `r q5_vec[3]`
  \item[d)] `r q5_vec[4]`
  \item[e)] `r q5_vec[5]`
\end{enumerate}

`r if (params$key) answer(q5_soln, q5_vec)`

*****

\newpage

```{r echo=FALSE, results='hide'}
prior <- sample(seq(.01, .08, .01), size = 1)
pos_given_sick <- sample(seq(.92, .98, .01), size = 1)
pos_given_healthy <- sample(seq(.04, .1, .01), size = 1)

pos <- pos_given_sick * prior + pos_given_healthy * (1 - prior)
post <- pos_given_sick * prior / pos

q6_vec <- sample(c(
  pos,
  prior,
  pos_given_sick * prior,
  pos_given_healthy * (1 - prior),
  1
))

q7_vec <- sample(c(
  post,
  pos_given_sick,
  prior,
  pos_given_healthy,
  pos_given_sick * prior
))
```

`r prior*100`\% of a population is infected with HIV. For a certain HIV test there is
a `r pos_given_sick*100`\% chance the test will be positive given that the tested
individual is actually infected with HIV. If the person is not infected, there is a 
`r pos_given_healthy*100`\% chance the test will be positive. Answer the following
**two** questions.

## Question 6

What is the probability that the test is positive for a randomly selected person 
from the population?

\begin{enumerate}
  \item[a)] `r q6_vec[1]`
  \item[b)] `r q6_vec[2]`
  \item[c)] `r q6_vec[3]`
  \item[d)] `r q6_vec[4]`
  \item[e)] `r q6_vec[5]`
\end{enumerate}

`r if (params$key) answer(pos, q6_vec)`

## Question 7

If a randomly selected person from the population is tested positive, what is 
the probability that the person is actually infected?

\begin{enumerate}
  \item[a)] `r q7_vec[1]`
  \item[b)] `r q7_vec[2]`
  \item[c)] `r q7_vec[3]`
  \item[d)] `r q7_vec[4]`
  \item[e)] `r q7_vec[5]`
\end{enumerate}

`r if (params$key) answer(post, q7_vec)`

*****

\newpage

```{r binomial, echo=FALSE, results='hide'}
prob <- sample(seq(.2, .8, .1), size = 1)

q9_ans <- paste0("pbinom(q = 2, size = 4, prob = ", prob, ")")
q9_vec <- sample(c(
  q9_ans,
  paste0("dbinom(x = 2, size = 4, prob = ", prob, ")"),
  paste0("pbinom(q = 2, size = 3, prob = ", prob+.2, ")"),
  paste0("rbinom(n = 2, size = 4, prob = ", prob, ")"),
  paste0("pgeom(q = 2, prob = ", prob, ")")
))

q10_ans <- 1 - (1-prob)^4
q10_vec <- sample(c(
  q10_ans,
  (1-prob)^4,
  dbinom(1, 4, prob),
  dbinom(2, 4, prob),
  sum(dbinom(1:3, 4, prob))
))
```

Let $X \sim Binomial(n = 4, p = `r prob`)$. Answer the following **three** questions.

## Question 8

Which of the following plots is of the pmf for $X$? The $y$ axis is $f(x)$ for each plot.

```{r echo=FALSE, fig.height = 5}
par(mfrow = c(3, 2), oma = c(5,4,0,0) + 0.1, mar = c(0,0,4,3) + 0.1)
plot(0:4, dbinom(0:4, 4, prob), type = "h", lwd = 2, ylab = "f(x)", xlab = "x") 
title(main = "Plot 1", line = 0.1)
plot(0:8, dbinom(0:8, 8, prob), type = "h", lwd = 2, ylab = "f(x)", xlab = "x")
title(main = "Plot 2", line = 0.1)
plot(0:4, rep(1/5, 5), type = "h", lwd = 2, ylab = "f(x)", xlab = "x") 
title(main = "Plot 3", line = 0.1)
plot(0:4, 3*dbinom(0:4, 4, prob), type = "h", lwd = 2, ylab = "f(x)", xlab = "x")
title(main = "Plot 4", line = 0.1)
plot(0:3, dbinom(0:3, 3, prob), type = "h", lwd = 2, ylab = "f(x)", xlab = "x") 
title(main = "Plot 5", line = 0.1)
```

\begin{enumerate}
  \item[a)] Plot 1
  \item[b)] Plot 2
  \item[c)] Plot 3
  \item[d)] Plot 4
  \item[e)] Plot 5
\end{enumerate}

`r if (params$key) paste0("SOLUTION: a")`

## Question 9

Which of the following correctly computes $P(X \le 2)$ in R?

\begin{enumerate}
  \item[a)] `r q9_vec[1]`
  \item[b)] `r q9_vec[2]`
  \item[c)] `r q9_vec[3]`
  \item[d)] `r q9_vec[4]`
  \item[e)] `r q9_vec[5]`
\end{enumerate}

`r if (params$key) answer(q9_ans, q9_vec)`

## Question 10

What is $P(X \ge 1)$?

\begin{enumerate}
  \item[a)] `r q10_vec[1]`
  \item[b)] `r q10_vec[2]`
  \item[c)] `r q10_vec[3]`
  \item[d)] `r q10_vec[4]`
  \item[e)] `r q10_vec[5]`
\end{enumerate}

`r if (params$key) answer(q10_ans, q10_vec)`

*****

Let $X \sim Binomial(N, p)$, where $N \sim Poisson(\lambda)$. Note that $N$ is a
random variable, not a number. Answer the following **two** questions.

```{r law_of_tot_expec, echo=FALSE}
q11_ans <- "$np$"
q11_vec <- sample(c(
  q11_ans,
  "$Np$",
  "$p$",
  "$n$",
  "$N$"
))

q12_ans <- "$ \\lambda p$"
q12_vec <- sample(c(
  q12_ans,
  "$np$",
  "$Np$",
  "$ \\lambda$",
  "$n \\lambda$"
))
```

## Question 11

What is $E[X|N = n]$? 

*Hint: read this as, if $N$ were not a random variable but an integer $n$, then
 what is the mean of $X$?*
 
\begin{enumerate}
  \item[a)] `r q11_vec[1]`
  \item[b)] `r q11_vec[2]`
  \item[c)] `r q11_vec[3]`
  \item[d)] `r q11_vec[4]`
  \item[e)] `r q11_vec[5]`
\end{enumerate}

`r if (params$key) answer(q11_ans, q11_vec)`

## Question 12

What is $E[X]$?

*Hint: use the law of total expectation. For random variables $A$ and $B$, 
$E[A] = \sum_b E[A|B=b]P(B=b)$.*

\begin{enumerate}
  \item[a)] `r q12_vec[1]`
  \item[b)] `r q12_vec[2]`
  \item[c)] `r q12_vec[3]`
  \item[d)] `r q12_vec[4]`
  \item[e)] `r q12_vec[5]`
\end{enumerate}

`r if (params$key) answer(q12_ans, q12_vec)`

*****

\newpage

```{r joint table, echo=FALSE, results='hide'}
table_prob <- sample(c(.1, .1, .2, .4, .1, .1))

y_pmf <- c(table_prob[1] + table_prob[4],
           table_prob[2] + table_prob[5],
           table_prob[3] + table_prob[6])

q13_denom <- sum(table_prob[1:3])
q13_ans <- table_prob[1:3] / q13_denom
q13_vec <- sample(list(
  q13_ans,
  table_prob[1:3],
  table_prob[4:6],
  table_prob[4:6] / sum(table_prob[4:6]),
  y_pmf
))

q14_ans <- sum((table_prob[1:3] / q13_denom) * (-3:-1))
q14_vec <- sample(c(
  q14_ans,
  sum(table_prob[1:3] * (-3:-1)),
  sum(table_prob[4:6] * (-3:-1)),
  sum((table_prob[4:6] / q13_denom) * (-3:-1)),
  sum(y_pmf * (-3:-1))
))
```

Consider the following joint probability distribution:

|             | X = 0             | X = 1                   |
| :---------- |:-------------:    | :-----:                 |
| Y = -3      | `r table_prob[1]` | `r table_prob[4]`       |
| Y = -2      | `r table_prob[2]` | `r table_prob[5]`       |
| Y = -1      | `r table_prob[3]` | `r table_prob[6]`       |

Answer the following **two** questions.

## Question 13

What is the conditional distribution of $Y$ given $X=0$, $f_{Y|X}(y)$?

\begin{enumerate}
  \item[a)] 
    $$
    \begin{aligned}
        f_{Y|X}(y) &= \begin{cases}
          `r q13_vec[[1]][1]`, & y = -3 \\
          `r q13_vec[[1]][2]`, & y = -2  \\
          `r q13_vec[[1]][3]`, & y = -1  \\
          0, & \text{otherwise}
        \end{cases}
    \end{aligned}
    $$
  \item[b)] 
    $$
    \begin{aligned}
        f_{Y|X}(y) &= \begin{cases}
          `r q13_vec[[2]][1]`, & y = -3 \\
          `r q13_vec[[2]][2]`, & y = -2  \\
          `r q13_vec[[2]][3]`, & y = -1  \\
          0, & \text{otherwise}
        \end{cases}
    \end{aligned}
    $$
  \item[c)] 
    $$
    \begin{aligned}
        f_{Y|X}(y) &= \begin{cases}
          `r q13_vec[[3]][1]`, & y = -3 \\
          `r q13_vec[[3]][2]`, & y = -2  \\
          `r q13_vec[[3]][3]`, & y = -1  \\
          0, & \text{otherwise}
        \end{cases}
    \end{aligned}
    $$
  \item[d)]
    $$
    \begin{aligned}
        f_{Y|X}(y) &= \begin{cases}
          `r q13_vec[[4]][1]`, & y = -3 \\
          `r q13_vec[[4]][2]`, & y = -2  \\
          `r q13_vec[[4]][3]`, & y = -1  \\
          0, & \text{otherwise}
        \end{cases}
    \end{aligned}
    $$
  \item[e)]
    $$
    \begin{aligned}
        f_{Y|X}(y) &= \begin{cases}
          `r q13_vec[[5]][1]`, & y = -3 \\
          `r q13_vec[[5]][2]`, & y = -2  \\
          `r q13_vec[[5]][3]`, & y = -1  \\
          0, & \text{otherwise}
        \end{cases}
    \end{aligned}
    $$
\end{enumerate}

`r if (params$key) answer_list(q13_ans, q13_vec)`

\newpage

## Question 14

What is $E[Y|X=0]?$

\begin{enumerate}
  \item[a)] `r q14_vec[1]`
  \item[b)] `r q14_vec[2]`
  \item[c)] `r q14_vec[3]`
  \item[d)] `r q14_vec[4]`
  \item[e)] `r q14_vec[5]`
\end{enumerate}

`r if (params$key) answer(q14_ans, q14_vec)`

*****

```{r likelihood, echo=FALSE, results='hide'}
breakpoint <- sample(-2:2, size = 1)
normmean <- breakpoint - 2

q15_ans <- breakpoint
q15_vec <- sample(c(
  q15_ans,
  normmean,
  5,
  -5,
  breakpoint + 1
))
```

## Question 15

Suppose we have $X_1, \ldots, X_n \sim f_\rho$, an iid sample of random variables 
from a discrete distribution with pmf $f_\rho(x)$ and parameter $\rho$. We 
construct the likelihood function and plot it:

```{r likelihood plot, echo=FALSE}
xn_vals <- seq(-5, breakpoint, .01)
xe_vals <- seq(breakpoint, 5, .01)
plot(xn_vals, dnorm(xn_vals, normmean), type = 'l', xlim = c(-5, 5), ylim = c(0, 1.1),
     xaxp = c(-5, 5, 10),
     ylab = expression(paste("L(", rho, ")")), xlab = expression(rho))
lines(xe_vals, dexp(xe_vals - breakpoint))
points(breakpoint, dnorm(breakpoint, normmean), cex = 2)
points(breakpoint, dexp(xe_vals[1] - breakpoint), cex = 2, pch = 19)
```

What is the maximum likelihood estimate for $\rho$?

\begin{enumerate}
  \item[a)] `r q15_vec[1]`
  \item[b)] `r q15_vec[2]`
  \item[c)] `r q15_vec[3]`
  \item[d)] `r q15_vec[4]`
  \item[e)] `r q15_vec[5]`
\end{enumerate}

`r if (params$key) answer(q15_ans, q15_vec)`

