---
title: "Exam 1 Review"
author: "Patrick Ding"
date: "9/30/2018"
header-includes:
   - \usepackage{tikz}
   - \usetikzlibrary{shapes.geometric,arrows}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1)
```

## 1

```{r echo=FALSE, fig.height=6}
create_cdf <- function(n,p,main=""){
  x <- 0:n
  DD <- x
  PP <- dbinom(x,n,p)
  QQ <- cumsum(PP)
  ## find x range for plotting (a little wider than min/max DD)
    xlims <- range(DD)
    slack <- .2*(xlims[2] - xlims[1])
    xlims[1] <- xlims[1] - slack
    xlims[2] <- xlims[2] + slack
    ## make plot
    par(mar=c(4.5,4.5,2,1))
    plot(c(0,0),xlim=xlims,ylim=c(0,1),ylab="F(x)",col=0,xlab="x",cex.axis=1.5,cex.lab=1.5,main=main,cex.main=2)
    segments(c(DD[1]-slack,DD),c(0,QQ),c(DD,DD[length(DD)]+slack),c(0,QQ),lwd=2)
    points(DD,c(0,QQ[1:(length(QQ)-1)]),lwd=2, cex=1.5)
    points(DD,QQ,pch=19,lwd=2, cex=1.5)
    box(which="figure")
}
params <- as.matrix(expand.grid(c(3,5),c(.1,.5,.9)))
to.use <- sample(1:6,5)
params <- params[to.use,]
truth <- sample(1:5,1)
others <- setdiff(1:5,truth)

par(mfrow=c(3,2))
for(ii in 1:5){
  create_cdf(params[ii,1],params[ii,2],paste("Plot ",ii,sep=""))
}
```

Which of the above plots is the cumulative distribution function for 
$X \sim Binomial(n=`r params[truth,1]`,p=`r params[truth,2]`)$?

## 2

If $B\subset A$, then $P(A\cup B)$ is
\begin{enumerate}
  \item[a)] $P(A)$
  \item[b)] $P(A)-P(A\cap B)$
  \item[c)] $P(A|B)$
  \item[d)] 1
  \item[e)] 0
\end{enumerate}

## 3

A plane has 4 engines. Let $X$ represent the number of engines that
 are working at any point in time. Consider the probability model shown below for $X$:
\begin{center}
  \begin{tabular}{l ll ll l}
    $X$        &	0       &	1&	2&	3&	4\\
    \hline
    Probability&	0.23    &0.27    &   0.11&   0.05&    0.34\\
  \end{tabular}
\end{center}
What is the expected number of engines that are working at any point in time?  
\begin{enumerate}
  \item[a)] 0.4
  \item[b)] 0.5 
  \item[c)] 1.5
  \item[d)] 2 
  \item[e)] 2.23 
\end{enumerate}

## 4

On an exam, the probability that a particular student will make a mistake on
any single question is 0.25. If there are 8 questions, and the
student's answers are mutually independent of each other, then the
probability that the student makes exactly two mistakes is
\begin{enumerate}
\item[a)] $\sum_{x=0}^2{8\choose x}(.25)^x(.75)^{8-x}$
\item[b)] $.689$
\item[c)] $.311$
\item[d)] $\sum_{x=2}^8{8\choose x}(.25)^x(.75)^{8-x}$
\item[e)] 1, since 2 is the expected value in this case
\end{enumerate}

*****

The probability mass function of the random variable $X$ is:
\begin{center}
  \begin{tabular}{|l| l|}
    \hline
    $x$        &  $f_X$\\
    \hline
    0 & 0.2\\
    1 & 0.8\\
    \hline
  \end{tabular}
\end{center}
Use this table to answer the next \textbf{two questions}.

## 5

What is $E[X]$?
\begin{enumerate}
  \item[a)] 0.8
  \item[b)] 0.16
  \item[c)] 1.0
  \item[d)] 0.5
  \item[e)] 0.64
\end{enumerate}

## 6

What is $Var(X)$?
\begin{enumerate}
  \item[a)] 0.20
  \item[b)] 0.16
  \item[c)] 0.10
  \item[d)] 0.25
  \item[e)] 0.64
\end{enumerate}

*****

A test has 20 questions with 5 possible answers each. A student who
has not studied at all decides to randomly guess an answer for each
question (with equal probability for each of the five answer
choices). Answer the following \textbf{three questions}. 

## 7

What is the expected number of correct answers?
\begin{enumerate} 
\item[a)] 3
\item[b)] 4
\item[c)] 5
\item[d)] 6
\item[e)] 7
\end{enumerate}

## 8

What is the standard deviation of the number of correct answers?
\begin{enumerate} 
\item[a)] 1.8
\item[b)] 4
\item[c)] 3.2
\item[d)] 2
\item[e)] 10
\end{enumerate}

## 9

What is the probability that the student gets more than one answer correct on the test?
\begin{enumerate} 
\item[a)] 0.07
\item[b)] 0.25
\item[c)] 0.75
\item[d)] 0.93
\item[e)] 0.97
\end{enumerate}

*****

In a batch of voltage regulators, 60\% came from supplier 1, 30\% from supplier 2 and 10\% from supplier 3.
\begin{itemize}
  \item 95\% of regulators from supplier 1 work
  \item 80\% of regulators from supplier 2 work
  \item 65\% of regulators from supplier 3 work
\end{itemize}
Use this information to answer the following \textbf{two questions}.

## 10

What is the probability a randomly selected regulator from the batch works?
\begin{enumerate}
\item[a)] 0.667
\item[b)] 0.875
\item[c)] 0.615
\item[d)] 0.125
\item[e)] 0.960
\end{enumerate}

## 11

If a randomly selected regulator from the batch works, what is the probability it came from supplier 3?
\begin{enumerate}
\item[a)] 0.65
\item[b)] 0.10
\item[c)] 0.05
\item[d)] 0.50
\item[e)] 0.07  %%CCR
\end{enumerate}

*****

## 12

Consider randomly selecting a student at TAMU.  Let $A$ denote the event
that the student has a MasterCard and $B$ denote the event that the
student has a Visa card.  Suppose that $P(A)=0.5$, $P(B)=0.6$, and
$P(A\cap B)=0.4$.  The probability that a randomly selected student 
at TAMU has exactly one of the two types of cards is

\begin{enumerate}
\item[a)]  0.3
\item[b)]  0.7
\item[c)]  0.1
\item[d)]  0.4
\item[e)]  0.2
\end{enumerate}

## 13

A discrete rv $X$ has a pmf given by
$$
p(x)=\begin{cases} C(x+1),&$x=0,1,2,3$ \\
            0,&otherwise.\end{cases}
$$
The value of $C$ that makes $p(x)$ a pmf is
\begin{enumerate}
\item[a)]  ${10}$
\item[b)]  ${1\over 6}$
\item[c)]  $1 \over 10$
\item[d)]  ${6}$
\item[e)]  ${1 \over 30}$
\end{enumerate}

## 14

Suppose that $X$ is a discrete rv with $E(X)=5$ and $V(X)=10$.  Suppose
that $Y=5X-2$.  Then the mean and variance of $Y$ are
\begin{enumerate}
\item[a)]  $E(Y)=23,\ V(Y)=250$
\item[b)]  $E(Y)=23,\ V(Y)=15.811$
\item[c)]  $E(Y)=23,\ V(Y)=50$
\item[d)]  $E(Y)=23,\ V(Y)=246$
\item[e)]  $E(Y)=23,\ V(Y)=48$
\end{enumerate}

## 15

Suppose $X \sim Binomial(10, .72)$. Which of the following would compute 
$P(X = 9)$ in R?

\begin{enumerate}
\item[a)]  rbinom(n = 10, size = 10, prob = .72)
\item[b)]  pbinom(q = 9, size = 10, prob = .72)
\item[c)]  dbinom(x = 9, size = 9, prob = .72)
\item[d)]  dbinom(x = 10, size = 9, prob = .72)
\item[e)]  dbinom(x = 9, size = 10, prob = .72)
\end{enumerate}

## 16

Suppose $W \sim Binomial(10, .3)$. Which of the following would compute
$P(3 \le W < 8)$ in R?

\begin{enumerate}
\item[a)] pbinom(q = 8, size = 10, prob = .3) - pbinom(q = 3, size = 10, prob = .3)
\item[b)] dbinom(x = 3:7, size = 10, prob = .3)
\item[c)] pbinom(q = 7, size = 10, prob = .3) - pbinom(q = 2, size = 10, prob = .3)
\item[d)] pbinom(q = 7, size = 10, prob = .3)
\item[e)] sum(dbinom(x = 3:8, size = 10, prob = .3))
\end{enumerate}

*****

Consider $x_1, \ldots, x_n \sim iid \; geometric(\gamma)$, where the pmf is 
$$f_\gamma(x_i) = (1-\gamma)^{x_i-1}\gamma$$
and $x =  1, 2, \ldots$, $0<\gamma<1$. Answer the following \textbf{two questions}. 

## 17

You construct the likelihood function by multiplying the pmfs evaluated at the 
data together: $\prod_{i=1}^n f_\gamma(x_i)$. So the likelihood function is of 
the form: $L(\gamma) = (1-\gamma)^A\gamma^B$. What is $A$ and $B$?

\begin{enumerate}
\item[a)] $A = x_1, \quad B = 1$
\item[b)] $A = \sum_{i=1}^n x_i - n, \quad B = 1$
\item[c)] $A = \prod_{i=1}^n x_i, \quad B = n$
\item[d)] $A = \sum_{i=1}^n x_i - n, \quad B = n$
\item[e)] $A = \sum_{i=1}^n x_i - 1, \quad B = n$
\end{enumerate}

## 18

What is the maximum likelihood estimate of $p$?

\begin{enumerate}
\item[a)] $\hat{p} = \frac{A}{B}$
\item[b)] $\hat{p} = \frac{A}{B} + 1$ 
\item[c)] $\hat{p} = \big(\frac{A}{B} - 1\big)^{-1}$ 
\item[d)] $\hat{p} = \big(\frac{A}{B} + 1\big)^{-1}$ 
\item[e)] $\hat{p} = A + B$
\end{enumerate}

*****

Consider the following joint probability distribution:

|             | X = -2             | X = 2               |
| :---------- |:-------------:   | :-----:               |
| Y = 3      |  .3               |  .1                   |
| Y = 4      |  .4               | .2                     |

Answer the following \textbf{two questions}. 

## 19

What is the marginal PMF of $X$?

\begin{enumerate}
\item[a)] 
  $$f_X(x) = \begin{cases} .3 & x = -2 \\ .1 & x = 2 \end{cases}$$
\item[b)] 
  $$f_X(x) = \begin{cases} .4 & x = -2 \\ .2 & x = 2 \end{cases}$$
\item[c)] 
  $$f_X(x) = \begin{cases} .7 & x = -2 \\ .3 & x = 2 \end{cases}$$
\item[d)] 
  $$f_X(x) = \begin{cases} .3 & x = -2 \\ .7 & x = 2 \end{cases}$$
\item[e)] 
  $$f_X(x) = \begin{cases} .4 & x = -2 \\ .6 & x = 2 \end{cases}$$
\end{enumerate}

## 20

What is $E[Y|X = 2]$?

\begin{enumerate}
\item[a)] $\frac{1}{3}$
\item[b)] $\frac{11}{3}$ 
\item[c)] $\frac{2}{3}$ 
\item[d)] $\frac{25}{7}$ 
\item[e)] $\frac{1}{4}$
\end{enumerate}

## Solutions

1. Plot `r truth`
2. a: $P(A)$
3. d: 2
4. c: .311
5. a: 0.8
6. b: 0.16
7. b: 4. If $X$ is the number of correct answers (guesses), then $X \sim Binomial(n=20,p=1/5)$. Therefore, $E(X) = np = 4$
8. a: 1.8. $Var(X) = np(1-p) = 20(1/5)(4/5) = 3.2 \Rightarrow sd(X) = \sqrt{3.2} = 1.8$
9. d: 0.93. $P(X>1) = 1-\Big(P(X=0) + P(X=1)\Big) = 1-\Big(\binom{20}{0} (1/5)^0(4/5)^{20} + \binom{20}{1} (1/5)^1(4/5)^{19}\Big) = 0.93$
10. b: $0.875 = .6\times.95 + .3\times.8 + .1\times.65$
11. e: $0.07 = (.1\times .65) / .875$
12. a: 0.3
13. c: $1 \over 10$
14. a: $E(Y)=23,\ V(Y)=250$ 
15. e: dbinom(x = 9, n = 10, p = .72)
16. c: pbinom(q = 7, size = 10, prob = .3) - pbinom(q = 2, size = 10, prob = .3)
17. d: $A = \sum_{i=1}^n x_i - n, \quad B = n$
18. d: $\hat{p} = \big(\frac{A}{B} + 1\big)^{-1}$ 
19. c: $f_X(x) = \begin{cases} .7 & x = -2 \\ .3 & x = 2 \end{cases}$
20. b: $\frac{11}{3}$ 



