---
title: 'Data Summary'
author: STAT 211 - 509
date: 2018/09/11
output:
  # tufte::tufte_handout: default
  xaringan::moon_reader:
    css: ["default", "tamu", "tamu-fonts"]
  lib_dir: libs
  nature:
    highlightStyle: github
    highlightLines: true
    countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
dt <- knitr::opts_knit$get("rmarkdown.pandoc.to")
set.seed(1)

source("../slide_tools.R")
```

# Descriptive statistics

+ We have a sample of data, drawn from some distribution

+ How to compute numerical summaries of the data?

+ How to visualize the data?

`r ns(dt)`

## Variables

+ **Variable**: any characteristic or quantity to be measured on units in a study

+ **Categorical variable**: places a unit into one of several categories

+ **Quantitative variable**: takes on numerical values


+ **Univariate**: data with one variable
+ **Bivariate**: data with two variables
+ **Multivariate**: data with three or more variables

`r ns(dt)`

## Example: US cereal

```{r}
dat <- MASS::UScereal
str(MASS::UScereal)
```

`r ns(dt)`

# Summarizing categorical variable

+ **Frequency**: number of times a value occurs in data

+ **Relative frequency**: proportion of data that has a value

```{r}
freqs <- table(dat$mfr)
freqs
props <- freqs / nrow(dat)
props
sum(props)
```

`r ns(dt)`

## Bar chart

+ Compares frequencies

+ Unordered

```{r fig.height=3}
barplot(freqs, cex.axis = .7, cex.lab = .7, cex = .7)
```

`r ns(dt)`

# Summarizing quantitative variable

+ What is the typical value of the variable?

+ What is the spread of the variable?

`r ns(dt)`

## Histogram

+ **Histogram**: bar graph of binned data where height of bar above each bin 
  denotes frequency or relative frequency of values in the bin
  
```{r fig.height=3}
hist(dat$calories, cex.axis = .7, cex.lab = .7, cex.main = .7)
```

`r ns(dt)`

+ Need to choose number of bins among which we divide the $n$ data points

+ General rule: number of bins $\approx \sqrt{n}$

+ `breaks` argument in `hist()`. Can be a string that specifies a built in algorithm
  for binning. Good default is "FD", for Freedman-Diaconis rule

`r ns(dt)`

## Breaks example  
  
+ Data drawn from a distribution with two modes

```{r fig.height=3}
set.seed(1)
z <- rbinom(20000, 1, .5) + 1
means <- c(-1, 1)
x <- rnorm(20000, mean = means[z], .7)
par(mfrow = c(1, 2))
hist(x, main = "Sturges", cex.main = .7, cex.axis = .7, cex.lab = .7)
hist(x, breaks = "FD", main = "Freedman-Diaconis", cex.main = .7, cex.axis = .7, cex.lab = .7)
```

## The shape of quantitative data

+ **Symmetric** data is mirrored about each side of a center value

```{r echo=FALSE, fig.height=3}
hist(rnorm(1000), breaks = "FD", main = NULL, cex.axis = .7, cex.lab = .7)
```

`r ns(dt)`

+ **Skewed** data has one side much longer than the other

```{r echo=FALSE, fig.height=3}
hist(rchisq(1000, df = 5), breaks = "FD", main = "Right skewed data", 
     cex.main = .7, cex.axis = .7, cex.lab = .7)
```

`r ns(dt)`

+ The **mode** is the peak value of the distribution

+ **Multimodal** data has multiple modes

```{r echo=FALSE, fig.height=3}
hist(x, breaks = "FD", main = "Bimodal data", cex.main = .7, cex.axis = .7, cex.lab = .7)
```

`r ns(dt)`

+ **Outliers** are data points "far" from most other data

+ Determination of outliers is subjective

+ *Do not* remove outliers if you don't know for sure that the data is erroneous

```{r echo=FALSE, fig.height=3}
x <- c(rnorm(100), 10, -20)
hist(x, breaks = "FD", main = "Data with outliers", cex.main = .7, cex.axis = .7, cex.lab = .7)
```

`r ns(dt)`

# Summary statistics for quantitative data

`r ns(dt)`

## Measures of central tendency

+ **Sample median**: value separating lower 50% of data from upper 50% of sample
    - For finite set of numbers, the middle value
    - If even number of values, then mean of middle two numbers

+ **Sample mean**: Given sample values $x_1, \ldots, x_n$, 

  $$\bar{x} = \frac{1}{n}\sum_{i=1}^nx_i$$

```{r}
mean(dat$calories)
```

`r ns(dt)`

## Percentiles

+ **Percentile**: the $p$th percentile is the value such that $p \times 100\%$ 
  of sample data is below it and $(1-p)\times 100\%$ are above it.
  
  - **First quartile (Q1)** is 25th percentile
  - **Second quartile (Q2)** is 50th percentile
  - **Third quartile (Q3)** is 75th percentile
  
+ **Five-number summary**

```{r}
fivenum(dat$calories)
summary(dat$calories)
```

`r ns(dt)`

## Boxplot

+ Visualize the 5 number summary

+ In R: `boxplot()`

+ **Interquartile range**: $IQR = Q3 - Q1$

+ Outliers: values greater than $Q3 + IQR$ or less than $Q1 - IQR$ are represented
  with a point

```{r echo=FALSE, fig.height=3}
boxplot(dat$calories, horizontal = TRUE, main = "Calories data", cex = .7, 
        cex.main = .7, cex.axis = .7)
text(x = min(dat$calories), y = 1.15, labels = "Min", cex = .7)
text(x = 110, y = 1.25, labels = "Q1", cex = .7)
text(x = 180, y = 1.25, labels = "Q3", cex = .7)
text(x = 135, y = 1.25, labels = "Q2", cex = .7)
text(x = 400, y = 1.15, labels = "Outliers", cex = .7)
```

`r ns(dt)`

## Measures of spread

- **IQR**: $Q3 - Q1$, the range of the middle 50% of the data

- **Sample variance, $s^2$**: sum of squared deviations from the mean divided by $n-1$:

  $$s^2 = \frac{1}{n-1}\sum_{i=1}^n(x_i - \bar{x})^2$$
  
- **Sample standard deviation, s**: square root of sample variance. Has same units
  as data
  
```{r}
var(dat$calories)
sd(dat$calories)
```

`r ns(dt)`

## Choosing measure of central tendency and spread

+ Sample mean and sample standard deviation good for symmetric data

+ For skewed data or data with outliers, sample median and interquartile range
  may be more appropriate

`r ns(dt)`

```{r echo=FALSE, fig.height=3}
set.seed(1)
x <- rnorm(100)
plot(density(x), cex.main = .7, cex.axis = .7, cex.lab = .7, main = "")
abline(v = mean(x), lwd = 3)
abline(v = mean(x) + sd(x), lwd = 3, lty = 3)
abline(v = mean(x) - sd(x), lwd = 3, lty = 3)
legend("topleft", legend = c("Mean", "plus/minus 1 sd"), lty = c(1, 3), lwd = 3,
       cex = .5)
```

```{r echo=FALSE, fig.height=3}
set.seed(1)
x <- c(rexp(100, rate = 1), 15:25)
plot(density(x), cex.main = .7, cex.axis = .7, cex.lab = .7, main = "")
abline(v = mean(x), lwd = 3, lty = 1)
abline(v = median(x), lwd = 3, lty = 3)
legend("topright", legend = c("Mean", "Median"), lty = c(1, 3), lwd = 3, cex = .5)
```
